---
title: "Using scatterbar with a SpatialExperiment object"
author: "Jean Fan and Dee Velazquez"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using scatterbar with a SpatialExperiment object}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# Using `scatterbar` with a `SpatialExperiment` object

This tutorial demonstrates how to visualize cell-type proportions with `scatterbar` from a `SpatialExperiment` object. `SpatialExperiment` is a class from Bioconductor that stores information from spatial-omics experiments, which we can use to visualize the cell types found in certain spots. We will use `SEraster` to rasterize cell-type counts and calculate their proportions within pixels, when can then be utilized by `scatterbar`.

For more information on `SpatialExperiment`, click [here] (https://www.bioconductor.org/packages/release/bioc/vignettes/SpatialExperiment/inst/doc/SpatialExperiment.html).

## Load libraries

First, we need to loading the necessary libraries and load in the dataset provided by `SEraster`. It is a preprocessed MERFISH dataset of the mouse preoptic area (POA) from a female naive animal. For more information, please refer to the original work, [Moffitt J. and Bambah-Mukku D. et al. (2018), "Molecular, spatial, and functional single-cell profiling of the hypothalamic preoptic region", *Science Advances*](https://www.science.org/doi/10.1126/science.aau5324).

```{r loading}
# Load required libraries
library(SpatialExperiment)
library(SEraster)
library(scatterbar)
library(ggplot2)

# Load the MERFISH dataset from mouse POA (Preoptic Area)
data("merfish_mousePOA")
```

We can see that this data is in the form of a `SpatialExperiment`.

```{r check}
# Check the class of the dataset
class(merfish_mousePOA)
```
## Rasterize Cell-Type Counts 

To aggregate cell-type data into spatial pixels, we use the `rasterizeCellType` function from `SEraster`. This function takes the `SpatialExperiment` object and generates a rasterized view of cell-type counts. We will rasterize at a resolution of 55 micrometers (µm) and use the "sum" function to aggregate the number of cells.

```{r rasterize}
# Rasterize the cell-type data at 55um resolution
rastCt <- SEraster::rasterizeCellType(
  merfish_mousePOA,  # SpatialExperiment object
  col_name = "celltype",  # Column with cell-type information
  resolution = 55,  # Set resolution to 55 micrometers
  fun = "sum",  # Sum up the cells within each pixel
  square = TRUE  # Use square-shaped pixels for rasterization
)

# Visualize the rasterized result (total number of cells per pixel)
SEraster::plotRaster(rastCt, name = "Total cells")
```

## Calculate Cell-Type Proportions

Next, we calculate the proportions of each cell type within each pixel. We first retrieve the list of cell IDs for each pixel and the corresponding cell types. We then calculate the proportions for each cell type by dividing the number of cells of a given type by the total number of cells in each pixel.

```{r prop}
# Extract the list of cell IDs for each pixel
cellids_perpixel <- colData(rastCt)$cellID_list

# Retrieve the cell-type information
ct <- merfish_mousePOA$celltype
names(ct) <- colnames(merfish_mousePOA)
ct <- as.factor(ct)  # Ensure cell types are factors

# Calculate proportions for each pixel
prop <- do.call(rbind, lapply(cellids_perpixel, function(x) {
  table(ct[x]) / length(x)
}))

# Set rownames to match the pixel IDs in the raster object
rownames(prop) <- rownames(colData(rastCt))
head(prop)  # Display the first few rows of the proportions matrix
```
## Retrieve Pixel Coordinates

For `scatterbar`, we also need the x and y coordinates of the pixels from the `rastCt` object. These spatial coordinates correspond to the positions of each pixel in the rasterized grid.

```{r pos}
# Extract the spatial coordinates of the pixels (x, y)
pos <- spatialCoords(rastCt)
head(pos)  # Display the first few rows of the spatial coordinates
```
## Filter Pixels with More than One Cell

We only want to visualize pixels that contain more than one cell, so we filter out pixels that do not meet this criterion.

```{r filter}
# Filter for pixels that only contain more than one cell for visualization
vi <- colData(rastCt)$num_cell > 1
pos <- pos[vi, ]  # Filter spatial coordinates
prop <- prop[vi, ]  # Filter proportions matrix

# Check dimensions to ensure filtering was successful
dim(pos)
dim(prop)
```
## Visualize Cell-Type Proportions Using `scatterbar`

Now that we have both the cell-type proportions and pixel position coordinates, we can visualize the data using `scatterbar`. We pass the proportions and coordinates, along with custom colors, to create a `scatterbar` plot. Remember that both the proportions and position data must be data frames in order to be passed into scatterbar.

```{r scatterbar}
# Generate custom colors for the cell types
custom_colors <- sample(rainbow(length(levels(ct))))

# Visualize the cell-type proportions using scatterbar
start.time <- Sys.time()
scatterbar::scatterbar(
  prop,  # Proportions matrix
  data.frame(pos),  # Spatial coordinates
  colors = custom_colors,  # Custom colors for each cell type
  padding_x = 10,  # Add padding to the x-axis
  padding_y = 10, # Add padding to the y-axis
  legend_title = "Cell Types" # Legend title
  ) + coord_fixed()  # Maintain aspect ratio
end.time <- Sys.time()
print(end.time - start.time)

```

This plot shows the proportion of each cell type within each pixel, with bars stacked to represent the composition of cell types. The colors correspond to different cell types, as defined by the custom color vector.
