---
title: "Using scatterbar with Visium data"
author: "Jean Fan and Dee Velazquez"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using scatterbar with Visium data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# Using scatterbar with Visium data

Below is how to use `scatterbar` from the provided Visium dataset of an FFPE preserved adult mouse brain partial coronal section from 10X Genomics.
```{r setup}
library(scatterbar)
library(ggplot2)
data("adult_mouse_brain_ffpe")

plot(adult_mouse_brain_ffpe$pos)
head(adult_mouse_brain_ffpe$prop)

start.time <- Sys.time()
scatterbar(
  adult_mouse_brain_ffpe$prop,
  adult_mouse_brain_ffpe$pos,
  size_x = 220,
  size_y = 220,
  legend_title = "Cell Types"
) + coord_fixed()

end.time <- Sys.time()
print(end.time - start.time)
```

Just like with the mOB data, we can change the order of how each bar is laid out by changing the order of the cell-type proportion matrix and combine `scatterbar` with other ggplot geoms and customization.
```{r shiftorder}
start.time <- Sys.time()
custom_colors <- c('1'= '#5d6f99',
    '2' = '#985a39',
    '3' =  '#d6589a',
    '4' = '#4d1395',
    '5' = '#b5ef27',
    '6' = '#77d5bc',
    '7' = '#7830d2',
    '8' ='#b43b59',
    '9' = '#1c40b1',
    '10' = "#FF5733",
    '11' = '#FFFF00',
    '12' = '#f4a6f1')
scatterbar::scatterbar(adult_mouse_brain_ffpe$prop[, c(2,3,4,11,5,6,10,7,8,1,9, 12)], adult_mouse_brain_ffpe$pos, size_x = 220, size_y = 220, padding_x = 0.1, padding_y = 0.1, legend_title = 'Cell Type', colors = custom_colors) +
  geom_point(data=adult_mouse_brain_ffpe$pos, mapping=aes(x=x, y=y), size = 0.1) +
  theme_bw() + ylab('y') + ggplot2::coord_fixed()
end.time <- Sys.time()
print(end.time - start.time)
```
