---
title: "Visualizing multiple datasets with scatterbar"
author: "Jean Fan and Dee Velazquez"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualizing multiple datasets with scatterbar}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# Visualizing multiple datasets with scatterbar

This tutorial demonstrates how to visualize multiple datasets together, utilizing the package `patchwork`.
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Below we can load in our datasets provided by `scatterbar` and create the respective scatterbars using those datasets and save them to a variable.
```{r setup}
library(scatterbar)
library(ggplot2)

data("mOB")
data("adult_mouse_brain_ffpe")

# Basic scatterbar plot with default settings
p1 <- scatterbar(mOB$data, mOB$xy) + coord_fixed()
p2 <- scatterbar(adult_mouse_brain_ffpe$prop, adult_mouse_brain_ffpe$pos) + coord_fixed()
```

We can then load in `patchwork` and visualize both scatterbars in one plot.
```{r patchwork}
library(patchwork)
p1 + p2

```
